\chapter{Full Wave Rectifier}
\vspace{-1cm}

%--------------------------AIM-----------------------------
\section{Aim}
\hspace{1cm}Single Phase Half Wave Uncontrolled and Controlled Rectifier

%--------------------------SOFTWARE USED-----------------------------
\section{Software Used}
\hspace{1cm}MATLAB R2020a
%-------------------------THEORY---------------------------
\section{Theory}
\hspace{\parindent}

A rectifier is an electronic device that converts AC (alternating current) to DC (direct current). This is typically achieved using diodes, which allow current to flow in only one direction. In the case of a half-wave rectifier, only one half-cycle of an AC voltage waveform is permitted to pass, while the other half-cycle is blocked. By using a single diode, half-wave rectifiers are able to convert AC voltage to DC voltage.

There are two main types of single-phase half-wave rectifiers: uncontrolled and controlled. A single-phase half-wave uncontrolled rectifier is made up of an AC source, a diode, and a load. During the positive half-cycle of the AC source, the diode becomes forward biased and the circuit conducts, while during the negative half-cycle, the diode becomes reverse biased and blocks current. This type of rectifier is simple and inexpensive to construct, but it has limited applications due to its inability to control the DC output.

A single-phase half-wave controlled rectifier, on the other hand, includes a thyristor or SCR (Silicon Controlled Rectifier) in addition to an AC source and a load. The key difference between a controlled rectifier and an uncontrolled rectifier is the presence of the thyristor/SCR, which conducts only when gate pulses at a firing angle $ \alpha $ are applied to it. Once the thyristor/SCR conducts, it continues to conduct even if the voltage across it becomes negative until the load current falls below the holding current level. The SCR automatically turns off when its voltage is reverse biased for a period longer than the SCR turn off time. This type of rectifier allows for better control of the DC output and is used in many applications, including power supplies, battery chargers, and DC motor drives.

In summary, rectifiers are important electronic devices that convert AC to DC voltage, and half-wave rectifiers are a type of rectifier that allows only one half-cycle of an AC voltage waveform to pass. While single-phase half-wave uncontrolled rectifiers are simple and inexpensive, they have limited applications due to their inability to control the DC output. Single-phase half-wave controlled rectifiers, which include a thyristor/SCR, allow for better control of the DC output and are used in many applications.


% \vspace{1cm}

%----------------------Theoretical Calculations----------------------
\section{Theoretical Calculations}
\hspace{\parindent}
The theoretical calculations for a half-wave rectifier with an R load are given by the formulas:


$$
    V_{o,avg} =
    V_{phase}
    \sqrt{2(1 + cos\alpha)2\pi} =
    V_m(1 + cos\alpha)
    2\pi
$$
$$
    I_{o,avg} =
    V_oR
$$

In uncontrolled rectifiers, $ \alpha$ = 0, and the thyristor is replaced with a diode.
For a single-phase half-wave uncontrolled rectifier with an RMS voltage of 230V
and a resistive load of 10$ \Omega $,
the output voltage is 103.53V, and the output current is 10.53A.

For a single-phase half-wave controlled rectifier with an RMS voltage of 230V
and a resistive load of 10$ \Omega $ and a firing angle of $ \alpha  $ = 30◦, the output voltage
is 96.6V, and the output current is 9.66A.

\pagebreak

% \input{chapters/chapter_1/circuits/circuit_1}
% \input{chapters/chapter_1/circuits/circuit_2}
% \input{chapters/chapter_1/circuits/circuit_3}
% \input{chapters/chapter_1/circuits/circuit_4}
% \input{chapters/chapter_1/circuits/circuit_5}
% \input{chapters/chapter_1/circuits/circuit_6}
% \input{chapters/chapter_1/circuits/circuit_7}

%-------------------------RESULTS---------------------------
\section{Results}
%------------SUB----------RESULTS---------------------------
\subsection{Theoretical Calculation}


\hspace{1.5cm} The resistance \textbf{R} in the circuit shown in Fig. 2.2 is considered as 20 $\Omega$, and the voltage (which is effectively the voltage across \textbf{R}) is considered as 50 V, 100 V, 150 V and 200 V in four steps as mentioned in the second column of Table-\ref{Table_simulation-result}. As per \textbf{Ohm's Law}, the current corresponding to all the four voltages are,
$$I=\frac{V}{R}=\frac{50}{20}= 2.5~A~~ (for~ V =50 V);~~~   = 5~A~~ (for~ V =100 V);~~~
    = 7.5~A~~ (for~ V =150 V);~~~= 10~A~~ (for~ V =200 V) $$


%------------SUB----------RESULTS---------------------------
\subsection{Simulation Results}
\hspace{1.5cm} The simulink file is run for 10 sec considering V=50 V, and and corresponding current seen in the display is noted in the fourth column of second row of Table-\ref{Table_simulation-result}. Similarly, all other three rows are filled.  Further, constantly varying ramp voltage is applied and the corresponding \textit{v-i} graph is plotted in Fig. \ref{Fig_v-i graph}.




\begin{table}[h]
    \renewcommand{\arraystretch}{1.3}
    \caption{Observations for Single Phase Half Wave Controlled Rectifier with RLE load}
    \centering
    \begin{tabular}{|c|c|c|c|}
        \hline
        \multirow{2}{*}{Sl No} & \multirow{2}{*}{Applied Voltage (V) in Volts} & \multicolumn{2}{c|}{Current (I) through R in Amps}             \\
        \cline{3-4}
                               &                                               & Theoretical                                        & Simulated \\
        \hline
        1                      & 50                                            & 5                                                  & 2.5       \\
        \hline
        2                      & 100                                           & 10                                                 & 5         \\
        \hline
        3                      & 150                                           & 15                                                 & 7.5       \\
        \hline
        4                      & 200                                           & 20                                                 & 10        \\
        \hline
    \end{tabular}
\end{table}


\section{Conclusion}
\hspace{\parindent}
The design of single phase half wave rectifiers, both controlled and
uncontrolled, with R, RL, RL with freewheeling diode, and RLE
loads were implemented successfully in MATLAB using Simulink.
The output waveforms for voltage and current were obtained in each
case, and a comparison between theoretically calculated and simu-
lated output parameter values was also performed.

\pagebreak